Fast Sorting on Modern Computers

### 2005-08-14 docs/tech_report/conclusions.tex
capitalization fix
done spell check. some errors found
removed the word project, and resolved the occasional TODO

\section{Results summary}

@@

However, we disagree with his results in several key places. Firstly, we note
that an in-lined sorting method does not perform as fast as insertion sort. We
find it difficult to perform his improvements on either algorithm N or algorithm
S, and write our own mergesort instead. We expected that mergesort would have
half the instruction count of heapsortl we found instead they were very similar.

@@

\section{Important results and contributions}

@@

\section{Instruction count results}

@@

% TODO more details in this
The test results in this project mostly agreed with LaMarca's results.  The
heapsort graphs had nearly exactly the same shape as LaMarca's, and the ratios
between the base and cache-conscious versions were nearly identical. The
multi-mergesort results were slightly different: instruction count was
identical, but while the cache results of the base and tiled mergesorts were
identical, multi-mergesort results were not the same. However, the basis for
those results - a sharp increase in cache misses followed by a flat-lined graph
- were in both cases.

@@

\afterpage{
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-10em}
\setcounter{figure}{9} % note the cheating to make this work
\begin{figure}[H]
\begin{changemargin}
\includegraphics[angle=90,scale=1.8]{plots/all_cycles}
\vspace*{8em}
\end{changemargin}
\caption[Cycles per key of several major sorts and their variations]{\label{All
cycles}Cycles per key of several major sorts and their variations - this was
measured on a Pentium 4 using hardware performance counters.}
\end{figure}
\newpage
}

@@

This section lists the contributions of this project. Each of these is novel  
and was discovered or developed during the course of this project.

The primary contribution of this project is that it repeats the work of LaMarca
on cache-conscious sorting, and validates many of his claims, while questioning
others. The next most important contribution is having performed a branch
prediction analysis of major sorts - which to my knowledge has not been done
before - and developing a framework by which these results can be repeated and
expanded upon.

>>>>

\section{Results Summary}

@@

However, we disagree with his results in several key places. Firstly, we note
that an in-lined sorting method does not perform as fast as insertion sort. We
find it difficult to perform his improvements on either algorithm N or algorithm
S, and write our own mergesort instead. We expected that mergesort would have
half the instruction count of heapsort we found instead they were very similar.

@@

\section{Important Results and Contributions}

@@

\section{Instruction Count Results}

@@

% TODO more details in this
The test results in this report mostly agreed with LaMarca's results.  The
heapsort graphs had nearly exactly the same shape as LaMarca's, and the ratios
between the base and cache-conscious versions were nearly identical. The
multi-mergesort results were slightly different: instruction count was
identical, but while the cache results of the base and tiled mergesorts were
identical, multi-mergesort results were not the same. However, the basis for
those results - a sharp increase in cache misses followed by a flat-lined graph
- were in both cases.

@@

\afterpage{
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-10em}
%\setcounter{figure}{9} % note the cheating to make this work
\begin{figure}[H]
\begin{changemargin}
\includegraphics[angle=90,scale=1.8]{plots/all_cycles}
\vspace*{8em}
\end{changemargin}
\caption[Cycles per key of several major sorts and their variations]{\label{All
cycles}Cycles per key of several major sorts and their variations - this was
measured on a Pentium 4 using hardware performance counters.}
\end{figure}
\newpage
}

@@

This section lists the contributions of this report. Each of these is novel  
and was discovered or developed during the course of this report.
%TODO cant say that
The primary contribution of this report is that it repeats the work of LaMarca
on cache-conscious sorting, and validates many of his claims, while questioning
others. The next most important contribution is having performed a branch
prediction analysis of major sorts - which to my knowledge has not been done
before - and developing a framework by which these results can be repeated and
expanded upon.
